---
title: "Project 2"
author: "Cheyenne McKinley and Justin Koss"
date: "2/25/19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

```{r packages}
# Load packages


```

## Study Information

### Response to Comments
  - Read and discuss your paper with your group partner
    Paper was discussed offline between Cheyenne and Justin.
  - Complete the nr-reporting-summary.pdf checklist (open in Adobe Acrobat)
    1) statistical test used and whether they are one sided or two sided: not relevant
    2) description of any assumptions or corrections: missing
    3) full description of the statistcal parameters including central tendency or other basic           estimates: missing
    4) for null hypothesis testing: not relevant
    5) for bayesian analysis: not relevant
    6) for heirarchical and complex designs: not relevant
    They only did ChIP-on-Chip. They did not run ChIP-seq in this paper.
  

### 1.	Title 

Genome-wide analysis of estrogen receptor binding

### 2.	Authorship

Jason S Carroll1, Clifford A Meyer2,3, Jun Song2,3, Wei Li2,3, Timothy R Geistlinger1, Je ́roˆme Eeckhoute1, Alexander S Brodsky4, Erika Krasnickas Keeton1, Kirsten C Fertuck1, Giles F Hall5, Qianben Wang1, Stefan Bekiranov6,8, Victor Sementchenko6, Edward A Fox5, Pamela A Silver5,7, Thomas R Gingeras6,
X Shirley Liu2,3 & Myles Brown1


### 3. QRP
1. Unreported variables: "We report the relevant statistics for only those PSSM cutoffs with maximum changes with respect to control."
The authors did not provide reasons to exclude these variables. Instead they should have reported all of the variables, even if only in the supplemental files.

We have no evidence to support that other QRPs were done in this study, however, unless the authors were transparent about them, we would not know about them.

2. There is no evidence of HARKing in this paper. However, one should take this with a grain of salt as no evidence does not mean they did not do it just that they were not transparent about it if they did.

3. There is no evidence of P-hacking. However, one should take this with a grain of salt as no evidence does not mean they did not do it just that they were not transparent about it if they did.

### 4.	Research Questions 

*3.1.	Please list each research question included in this study.*

1. This study does not list research questions, but instead identifies a research objective, which is as follows, "... the identification of cis-regulatory sites in previously unexplored regions of the genome and the cooperating transcription factors underlying estrogen signaling in breast cancer."


### 5.	Hypotheses

*4.1.	For each of the research questions listed in the previous section, provide one or multiple specific and testable hypotheses. Please state if the hypotheses are directional or non-directional. If directional, state the direction. A predicted effect is also appropriate here.*

1. Since this study is observational, there was no testable hypothesis.
 
## Sampling Plan

In this section we will ask you to describe how you plan to collect samples, as well as the number of samples you plan to collect and your rationale for this decision. Please keep in mind that the data described in this section should be the actual data used for analysis, so if you are using a subset of a larger dataset, please describe the subset that will actually be used in your study.

### 5.	Existing data

*5.1.	Preregistration is designed to make clear the distinction between confirmatory tests, specified prior to seeing the data, and exploratory analyses conducted after observing the data. Therefore, creating a research plan in which existing data will be used presents unique challenges. Please select the description that best describes your situation. Please do not hesitate to contact us if you have questions about how to answer this question (prereg@cos.io).*

We will not be conducting preregistration, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.

- Registration prior to creation of data: As of the date of submission of this research plan for preregistration, the data have not yet been collected, created, or realized. 

- Registration prior to any human observation of the data: As of the date of submission, the data exist but have not yet been quantified, constructed, observed, or reported by anyone - including individuals that are not associated with the proposed study. Examples include museum specimens that have not been measured and data that have been collected by non-human collectors and are inaccessible.

- Registration prior to accessing the data: As of the date of submission, the data exist, but have not been accessed by you or your collaborators. Commonly, this includes data that has been collected by another researcher or institution.

-	Registration prior to analysis of the data: As of the date of submission, the data exist and you have accessed it, though no analysis has been conducted related to the research plan (including calculation of summary statistics). A common situation for this scenario when a large dataset exists that is used for many different studies over time, or when a data set is randomly split into a sample for exploratory analyses, and the other section of data is reserved for later confirmatory data analysis.

-	Registration following analysis of the data: As of the date of submission, you have accessed and analyzed some of the data relevant to the research plan. This includes preliminary analysis of variables, calculation of descriptive statistics, and observation of data distributions. Studies that fall into this category are ineligible for the Pre-Reg Challenge. Please contact us (prereg@cos.io) and we will be happy to help you.

### Data Required 
In order to confirm the analysis we would need the unaltered raw data that they used to conduct their statistical tests.


### 6.	Explanation of existing data

*6.1.	If you indicate that you will be using some data that already exist in this study, please describe the steps you have taken to assure that you are unaware of any patterns or summary statistics in the data. This may include an explanation of how access to the data has been limited, who has observed the data, or how you have avoided observing any analysis of the specific data you will use in your study. The purpose of this question is to assure that the line between confirmatory and exploratory analysis is clear.*  

We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.

### 7.	Data collection procedures.

*7.1.	Please describe the process by which you will collect your data. If you are using human subjects, this should include the population from which you obtain subjects, recruitment efforts, payment for participation, how subjects will be selected for eligibility from the initial pool (e.g. inclusion and exclusion rules), and your study timeline. For studies that don’t include human subjects, include information about how you will collect samples, duration of data gathering efforts, source or location of samples, or batch numbers you will use.* 

We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.

### 8.	Sample size

*8.1.	Describe the sample size of your study. How many units will be analyzed in the study? This could be the number of people, birds, classrooms, plots, interactions, or countries included. If the units are not individuals, then describe the size requirements for each unit. If you are using a clustered or multilevel design, how many units are you collecting at each level of the analysis?*

3665 estrogen receptor binding sites and 3629 RNA PolII binding sites

### 9.	Sample size rationale

*9.1.	This could include a power analysis or an arbitrary constraint such as time, money, or personnel.*

That is the amount identified after BLAT analysis.

### 10.	Stopping rule

*10.1.	If your data collection procedures do not give you full control over your exact sample size, specify how you will decide when to terminate your data collection.* 

 Not Applicable 

## Variables

In this section you can describe all variables (both manipulated and measured variables) that will later be used in your confirmatory analysis plan. In your analysis plan, you will have the opportunity to describe how each variable will be used. If you have variables that you are measuring for exploratory analyses, you are not required to list them, though you are permitted to do so.

### 11.	Manipulated variables

*11.1.	Describe all variables you plan to manipulate and the levels or treatment arms of each variable. For observational studies and meta-analyses, simply state that this is not applicable.*

Not applicable

### 12.	Measured variables

*12.1.	Describe each variable that you will measure. This will include outcome measures, as well as any predictors or covariates that you will measure. You do not need to include any variables that you plan on collecting if they are not going to be included in the confirmatory analyses of this study.*

RNA Pol density vs Distance from transcription start site, Conservation score vs Distance from center of binding sites (bp), number of binding sites vs number of genes, number of binding sites vs nucelotides, estrogen, number of motifs vs distance from center for ERE, AP*1, FKH, Oct, and C/EBP, MRNA level (as a multiple of siLUC control) vs time (h), transcript level (as a multiple of input) vs time (h), transcript level (as a multiple of vehicle) vs time (h)

### 13.	Indices

*13.1.	If any measurements are going to be combined into an index (or even a mean), what measures will you use and how will they be combined? Include either a formula or a precise description of your method. If you are using a more complicated statistical method to combine measures (e.g. a factor analysis), you can note that here but describe the exact method in the analysis plan section.*

Not applicable

## Design Plan

In this section, you will be asked to describe the overall design of your study. Remember that this research plan is designed to register a single study, so if you have multiple experimental designs, please complete a separate preregistration.



### 14.	Study type

-	Experiment - A researcher randomly assigns treatments to study subjects, this includes field or lab experiments. This is also known as an intervention experiment and includes randomized controlled trials.

[X]	Observational Study - Data is collected from study subjects that are not randomly assigned to a treatment. This includes surveys, natural experiments, and regression discontinuity designs.

-	Meta-Analysis - A systematic review of published studies.

-	Other - please explain.


### 15.	Blinding

*15.1.	Blinding describes who is aware of the experimental manipulations within a study. Mark all that apply.*

[X]	No blinding is involved in this study.

-	For studies that involve human subjects, they will not know the treatment group to which they have been assigned.

-	Personnel who interact directly with the study subjects (either human or non-human subjects) will not be aware of the assigned treatments.

-	Personnel who analyze the data collected from the study are not aware of the treatment applied to any given group.

### 16.	Study design

*16.1.	Describe your study design. Examples include two-group, factorial, randomized block, and repeated measures. Is it a between (unpaired), within-subject (paired), or mixed design? Describe any counterbalancing required. Typical study designs for observation studies include cohort, cross sectional, and case-control studies.*

Cross sectional

### 17.	Randomization

*17.1.	If you are doing a randomized study, how will you randomize, and at what level?*

We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.

## Analysis Plan

You may describe one or more confirmatory analysis in this preregistration. Please remember that all analyses specified below must be reported in the final article, and any additional analyses must be noted as exploratory or hypothesis generating.

A confirmatory analysis plan must state up front which variables are predictors (independent) and which are the outcomes (dependent), otherwise it is an exploratory analysis. You are allowed to describe any exploratory work here, but a clear confirmatory analysis is required. 

### 18.	Statistical models

*18.1.	What statistical model will you use to test each hypothesis? Please include the type of model (e.g. ANOVA, multiple regression, SEM, etc) and the specification of the model (this includes each variable that will be included as predictors, outcomes, or covariates). Please specify any interactions that will be tested and remember that any test not included here must be noted as an exploratory test in your final article.*

We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.

### 19.	Transformations

*19.1.	If you plan on transforming, centering, recoding the data, or will require a coding scheme for categorical variables, please describe that process.*

We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.


### 20.	Follow-up analyses

*20.1.	If not specified previously, will you be conducting any confirmatory analyses to follow up on effects in your statistical model, such as subgroup analyses, pairwise or complex contrasts, or follow-up tests from interactions. Remember that any analyses not specified in this research plan must be noted as exploratory.*

We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.

### 21.	Inference criteria

*21.1.	What criteria will you use to make inferences? Please describe the information you will use (e.g. p-values, Bayes factors, specific model fit indices), as well as cut-off criterion, where appropriate. Will you be using one or two tailed tests for each of your analyses? If you are comparing multiple conditions or testing multiple hypotheses, will you account for this?*

We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.

### 22.	Data exclusion

*22.1.	How will you determine what data or samples, if any, to exclude from your analyses? How will outliers be handled?*

We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.  

### 23.	Missing data

*23.1.	How will you deal with incomplete or missing data?*

We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.

### 24.	Exploratory analysis (optional)

*24.1.	If you plan to explore your data set to look for unexpected differences or relationships, you may describe those tests here. An exploratory test is any test where a prediction is not made up front, or there are multiple possible tests that you are going to use. A statistically significant finding in an exploratory test is a great way to form a new confirmatory hypothesis, which could be registered at a later time.* 


We cannot answer this question, because we will not be conducting a follow-up/replication study, due to unavailabiltiy of the data.

## Script (Optional) 

The purpose of a fully commented analysis script is to unambiguously provide the responses to all of the questions raised in the analysis section. This step is not common, but we encourage you to try to create an analysis script, refine it using a modeled dataset, and use it in place of your written analysis plan.

### 25.	Analysis scripts (Optional)

*25.1.	(Optional) Upload an analysis script with clear comments. This optional step is helpful in order to create a process that is completely transparent and increase the likelihood that your analysis can be replicated. We recommend that you run the code on a simulated dataset in order to check that it will run without errors.*

 

## Other

### 26.	Other

*26.1.	If there is any additional information that you feel needs to be included in your preregistration, please enter it here.*

In order to answer most of these questions and conduct a follow up study we need access to the data. The authors provide a website to download the data (linked below) however, the data seems to be corrupted and is unable to be opened/downloheaded. Until we would be able to obtain this data, there is no way to replicate the study. This highlights the problem with replicability, as this data was not publicly maintained, and on a private server. In the future, data should be made available on publicly maintained databases so that it can be available in the future, should something happen to the authors, personal technology, or the lab. This is something that Bucknell has done, and other institutions should also do the same.

To gain access to the data one of the researchers on the project had to be emailed and he then had to respond and personally send us the data. This also poses a problem in regards to the reproducibility of this study. First, the majority of the time the researchers do not respond to email requests asking for the data. Second, alot of times a researcher may have moved and their listed email changed so they become unreachable. Finally, while the researcher did send us the data, what he actually sent was the data embedded in another paper with a different title and due to the size of the data set it becomes quite difficult to ensure that the data is the exact same as the data we were originally looking for.

http://research.dfci.harvard.edu/brownlab/datasets/index.php? dir=ER_whole_human_genome/

### 27. Repeatability from other sources
Article: Repeatability of published microarray gene expression analyses
Authors: Ioannidis, John P A; Allison, David B; Ball, Catherine A; Coulibaly, Issa; Cui, Xiangqin; et al.
Link: https://search.proquest.com/docview/222636826/fulltextPDF/633BF533BF5A4604PQ/1?accountid=9784

This article tested the reproducibility of our and many other similar studies. While this article was able to access the data, this needs to be taken with a grain of salt as this article was published in 2009 (i.e. ten years ago). According to this article, our article did not state the software they used or the parameters for image processing. To test the reproducibility of our study, Ioannidis et al. specifically tried to recreate figures 2 a and b from our study and they determined that some aspects were reproducibale but overall there were discrepencies. While Ioannidis et al. was able to recreate the heat map, the number of significant genes they obtained was different then the published result. Ioannidis et al. also highlighted the fact that our study did not specify and state the specific p-value that they were using as their cutoff threshold. Finally, when Ioannidis et al. ran the statistical analysis that our paper did to generate figure 2b, the results that they got were different than the results that were published in our study. 

## References 

Due to the difficulty that can arise with two people on the same R script and trying to push that  script at the same time, Cheyenne and Justin verbally discussed the answers to all questions with only one person at a time typing on the actual R script. 